---
title: "Youtube Video Practice"
author: "Patrick Chase"
date: "1/25/2021"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Introduction to ggplot 
link- https://www.youtube.com/watch?v=UK2FhxMnmjQ 

```{r}

library(magrittr)
library(tidyverse)
library(ggplot2)
data(mpg)

head(mpg)

plot(mpg$displ, mpg$hwy)


```

ggplot work 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class))
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, size = class))

```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, alpha = class))
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  facet_wrap(~ class, nrow = 2) 
```
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_wrap(~ class, nrow = 2) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    caption = "Data from fueleconomy.gov",
    x="Engine displacement (liters)",
    y="Highway gas milage (mpg)"
  )
```

```{r}
p1 = ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

## adding facet later 

p1 + facet_wrap( ~ class, nrow = 1)
p1 + facet_wrap(~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue") 
```

Data workflow + bar plots 
link https://www.youtube.com/watch?v=k76R7ifcyvs 


```{r}
xtabs(~survived + passengerClass, data=titanic) %>% 
  prop.table(margin=2)
```
group/pipe/summarize

```{r}
titanic %>% 
  group_by(sex) %>%
  summarize(total_count = n(), 
            mean_age = mean(age),
            surv_count = sum(survived == 'yes'), 
            surv_pct = surv_count/total_count
            )
```
```{r}
titanic %>% 
  group_by(sex) %>%
  summarize (mean_age = mean(age), 
             surv_pct = sum(survived == 'yes')/n())
```
Filtering
```{r}
titanic %>% 
  filter(sex == 'female') %>% 
  group_by(passengerClass) %>%
  summarize(total_count = n(), 
            surv_pct = sum(survived == 'yes')/n())
```

```{r}
titanic %>% 
  filter(age < 18) %>% 
  group_by(passengerClass) %>%
  summarize(total_count = n(),
            surv_pct = sum(survived == 'yes')/n())
```

```{r}
titanic = titanic %>% 
  mutate(adult = ifelse(age > 18, '18+', 'under 18')) 
head(titanic)
```

```{r}
titanic %>% 
  group_by(sex, adult, passengerClass) %>% 
  summarize(surv_pct = sum(survived == 'yes')/n())

```

Bar Plots

Define data into a data frame d1
```{r}
d1 = titanic %>% 
  group_by(sex) %>% 
  summarize(surv_pct = sum(survived == 'yes')/n())
d1
```
now use d1 to create bar plot 
```{r}
ggplot(data = d1) +
  geom_col(mapping = aes(x=sex, y=surv_pct))

  
```

Grouping by 2 variables 
```{r}
d2 = titanic %>%
  group_by(sex, adult) %>%
  summarize(surv_pct = sum(survived == 'yes')/n())
d2
  
```

Now use d2 to generate a bar plot 
```{r}
ggplot(data = d2) +
  geom_col(mapping = aes(x=adult, y=surv_pct, fill=sex), 
           position = 'dodge')
```

Three conditioning variables
```{r}
d3 = titanic %>% 
  group_by(sex, passengerClass, adult) %>%
  summarize(surv_pct = sum(survived == 'yes')/n())

ggplot(data = d3) + 
  geom_col(mapping = aes(x=adult, y=surv_pct, fill=sex),
           position = 'dodge') +
  facet_wrap(~passengerClass)
```

3 conditioning variables with pretty labels

```{r}
d3 = titanic %>% 
  group_by(sex, passengerClass, adult) %>%
  summarize(surv_pct = sum(survived == 'yes')/n())

ggplot(data = d3) + 
  geom_col(mapping = aes(x=adult, y=surv_pct, fill=sex),
           position = 'dodge') +
  facet_wrap(~passengerClass) +
  labs(title = "Survival on the Titanic",
       y = "Fraction Surviving",
       x = "Age", 
       fill = "Sex")
```

Same information different organization
```{r}
ggplot(data = d3) + 
  geom_col(mapping = aes(x=sex, y=surv_pct, fill=passengerClass),
          position = 'dodge') +
  facet_wrap(~adult)
```
```{r}
d4 = titanic %>% 
  group_by(sex, passengerClass) %>%
  summarize(mean_age = mean(age))
d4
```

```{r}
ggplot(data = d4) +
  geom_col(mapping = aes(x=passengerClass, y=mean_age, fill=sex),
           position = 'dodge')
```

Visualizing Distributions
link https://www.youtube.com/watch?v=wFpzPtdIfTg 

```{r}
library(tidyverse)
library(ggplot2)

hist(us_counties$median_income2018)
```


histogram in ggplot2 
```{r}
ggplot(data = us_counties) + 
  geom_histogram(aes(x=median_income2018), binwidth = 2000)
```

density histogram 
```{r}
ggplot(data=us_counties) + 
  geom_histogram(aes(x=median_income2018, after_stat(density)), binwidth = 2000) 
```













